---
title: "DATA 605 HOMEWORK WEEK 1 - IMAGE MANIPULATION"
author: "Enid Roman"
date: "2023-01-29"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(dplyr)
library(plotly)
library(ggplot2)
#install.packages("gifski")
library(gifski)
#install.packages("anim.plots")
library(anim.plots)
```


**Build the first letters for both your first and last name using point plots in R.**


```{r}

x=c(seq(-1,-.5,length.out=500),
    rep(-1,500),
    seq(-1,-.5,length.out=500),
    seq(-1,-.5,length.out=500),
    seq(.5,1,length.out=500),
    seq(.5,1,length.out=500),
    rep(.5,500),
    rep(1,500),
    rep(.5,500),
    seq(1,.5,length.out = 500))
y=c(rep(-1,500),
    seq(-1,1,length.out=500),
    rep(-0,500),
    rep(1,500),
    rep(1,500),
    rep(0,500),
    seq(-1,1,length.out=500),
    seq(1,0,length.out=500),
    seq(0,-1,length.out=500),
    seq(-1,0,length.out = 500))
z=rbind(x,y)
plot(y~x, xlim=c(-3,3), ylim=c(-3,3),col='blue')

```

**Write R code that will left multiply (%>%) a square matrix (x) against each of the vectors of points (y). Initially, that square matrix will be the Identity matrix.**


```{r}

leftMultiply <- function(x,y){
   x %*% y
}
leftMultiply(matrix(rep(seq(1,3, length.out=3),3), nrow = 3, ncol = 3),diag(3))
```

**Use a loop that changes the transformation matrix incrementally to demonstrate 1) shear, 2) scaling, 3)  rotation , and 4) projection in animated fashion.**


**SHEAR**


```{r, animation.hook='gifski'}

for (i in seq(0,1,length.out=8)) {
  z1<-apply(z,2,function(x) leftMultiply(x,matrix(c(1,i,0,1), nrow=2, ncol=2)))
   plot(z1[2,]~z1[1,], xlim=c(-3,3), ylim=c(-3,3), col='blue')
}

```

**SCALING**


```{r, animation.hook='gifski'}

for (i in seq(0,1,length.out=10)) {
  z1<-apply(z,2,function(x) leftMultiply(x,matrix(c(i,0,0,i), nrow=2, ncol=2)))
   plot(z1[2,]~z1[1,], xlim=c(-3,3), ylim=c(-3,3), col='RED')
}
```

**ROTATION**


```{r, animation.hook='gifski'}

for (i in seq(0,pi*2,length.out=10)) {
  z1<-apply(z,2,function(x) leftMultiply(x,matrix(c(cos(i), -sin(i), sin(i), cos(i)), nrow=2, ncol=2)))
   plot(z1[2,]~z1[1,], xlim=c(-3,3), ylim=c(-3,3), col='YELLOW')
}
```

**PROJECTION**


```{r, animation.hook='gifski'}

for (i in seq(0,2*pi,length.out=12)) {
  tempZ<-rbind(z,rep(0, ncol(z)))
  z1<-apply(tempZ,2,function(x) leftMultiply(x,matrix(c(1,0,0,0, cos(i), -sin(i), 0 , sin(i), cos(i)), nrow=3, ncol=3)))
  plot(z1[2,]~z1[1,], xlim=c(-3,3), ylim=c(-3,3), col='DARK GREEN')
}
```






