---
title: "E.Roman_HW4.rmd"
author: "Enid Roman"
date: "2023-02-19"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read Pix Set Up Initialization

```{r}
library(jpeg)
library(OpenImageR)
library(EBImage)
```

```{r}
getwd()
```
**Use of Graphics**

We add graphics to the data set.

```{r}
# Prepare for Image Processing
num=17
files=list.files("./jpg/",pattern=".jpg")[1:num] 
```

```{r}
#Setting the file path
dim(readJPEG(paste0(file.path(getwd(),'jpg//'),files[1])))
```
```{r}
is.na(getwd())
```
**View Shoes Function**

```{r}
# Set Adj. Parameters
height=1200; width=2500;scale=20
plot_jpeg = function(path, add=FALSE)
{ jpg = readJPEG(path, native=T) # read the file
  res = dim(jpg)[2:1] # get the resolution, [x, y]
  if (!add) # initialize an empty plot area if add==FALSE
    plot(1,1,xlim=c(1,res[1]),ylim=c(1,res[2]),asp=1,type='n',xaxs='i',yaxs='i',xaxt='n',yaxt='n',xlab='',ylab='',bty='n')
  rasterImage(jpg,1,1,res[1],res[2])
}
```

**Load the Data into an Array**

This is loading the data in the array by each file, an image structure into the array
Dimensions will be number of images x ht/scale x wt/scale, channel.
There are 3 channels for red, green and blue.
Resize from the EBI package will scale the images to the desired dimension.

```{r}
# Load
im=array(rep(0,length(files)*height/scale*width/scale*3), dim=c(length(files), height/scale, width/scale,3))

for (i in 1:17){
temp=resize(readJPEG(paste0('jpg//', files[i])),height/scale, width/scale)
  im[i,,,]=array(temp,dim=c(1, height/scale, width/scale,3))}
```

**Vectorize**

The initializing array built in the previous step now needs to be modified into a 2 dimensional matrix.
The dimensions will be the number of images x (ht/s * wt/s * 3)
Essentially, reshaping the array into a vector. Dimension (number of image x(numberHt/sWt/s*3))

```{r}
flat=matrix(0, 17, prod(dim(im))) 
for (i in 1:17) {
  newim <- readJPEG(paste0('jpg//', files[i]))
  r=as.vector(im[i,,,1]); g=as.vector(im[i,,,2]);b=as.vector(im[i,,,3])
  flat[i,] <- t(c(r, g, b))
}
shoes=as.data.frame(t(flat))
dim(flat)
```
```{r}
dim(shoes)
```
```{r}
str(shoes)
```
**Actual Plots**

Using the plot_jpeg function to see the images

```{r}
#Old Shoes
par(mfrow=c(3,3))
par(mai=c(.3,.3,.3,.3))
for (i in 1:17){  #plot the first images only
plot_jpeg(writeJPEG(im[i,,,]))
}
```

**Get Eigencomponents from Correlation Structure**

We are scaling the pixels in the array.
Now we need to scale and center data using the scale function. We are normalizing the data to mean=0 and sd=1.

```{r}
scaled=scale(shoes, center = TRUE, scale = TRUE)
mean.shoe=attr(scaled, "scaled:center") #saving for classification
std.shoe=attr(scaled, "scaled:scale")  #saving for classification...later
dim(scaled)
```
```{r}
str(scaled)
```
**Calculate Covariance (Correlation)**

We are using the Correlation matrix. All the preceeding work was done to get to this matrix.
It is this correlation matrix that will be used to determine the eigencomponents.

```{r}
Sigma_=cor(scaled)
dim(Sigma_)
```
```{r}
str(Sigma_)
```
```{r}
Sigma_
```
**Get the Eigencomponents**

We use the Correlation matrix to find the eigenvectors and eigenvalues.
The matrix will find the mulitpliers (eigenvalues) to the vectors (eigenvectors) for the new linear combination so that the most variability is on the 1st component and descends thereafter.
This is a restructing that is used to reduce the dimensionality.
The eigenvalues are the proportion of variability associated with each component.
Specifically, the cumulative proportion of variability is sought after.

```{r}
myeigen=eigen(Sigma_)
myeigen
```
```{r}
cumsum(myeigen$values) / sum(myeigen$values)
```


**Eigenshoes**

The new shoes are a reduced number that are built by the new linear combinations. The new shoes are eigenvectors.

80% is between 2 and 3 components….. Here is 2

```{r}
scaling=diag(myeigen$values[1:2]^(-1/2)) / (sqrt(nrow(scaled)-1))

scaling
```
```{r}
eigenshoes=scaled%*%myeigen$vectors[,1:2]%*%scaling
dim(eigenshoes)
```
```{r}
str(eigenshoes)
```
Image 1

```{r}
scaling=diag(myeigen$values[1:5]^(-1/2)) / (sqrt(nrow(scaled)-1))
eigenshoes=scaled%*%myeigen$vectors[,1:5]%*%scaling
par(mfrow=c(2,3))
imageShow(array(eigenshoes[,1], c(60,125,3)))
```

Image 2

```{r}
imageShow(array(eigenshoes[,2], c(60,125,3)))
```

**Generate Principal Components**
Transform the images

```{r}
#Generate Variables
height=1200
width=2500
scale=20
newdata=im
dim(newdata)=c(length(files),height*width*3/scale^2)
mypca=princomp(t(as.matrix(newdata)), scores=TRUE, cor=TRUE)
################################################################
```

**Eigenshoes**
Generate Eigenshoes

```{r}
#Eigenshoes
mypca2=t(mypca$scores)
dim(mypca2)=c(length(files),height/scale,width/scale,3)
par(mfrow=c(5,5))
par(mai=c(.001,.001,.001,.001))
for (i in 1:17){#plot the first 25 Eigenshoes only
plot_jpeg(writeJPEG(mypca2[i,,,], bg="white"))  #complete without reduction
}
```
```{r}
a=round(mypca$sdev[1:17]^2/ sum(mypca$sdev^2),3)
cumsum(a)
```

**New Data Set**

```{r}
x = t(t(eigenshoes)%*%scaled)
```


**REFERENCES**
Fulton, Larry. Eigenshoes. https://rpubs.com/R-Minator/eigenshoes.







